Reconstructing the Lensing Mass in the Universe \\ from Photometric Catalogue Data

### 2012-11-20 doc/paper-lcr.tex
paper update

Along lines of sight containing strong gravitational lenses, the
perturbative effects of line of sight mass structure have been found to
be particularly important, with foreground and background structures
having a significant effect on the inferred lensing cross-section
\citep[\eg][]{WongEtal2012} and distance ratios
\citep[][]{DalalEtal2005}. Indeed, \citet{SuyuEtal2010} found that in
time delay lens cosmography the so-called ``external convergence''
$\kappax$ due to mass structures along the (unusually over-dense) line
of sight to the quadruply-imaged radio source B1608$+$656 was the
dominant source of uncertainty in their 5\% measurement of the time delay distance. 
Whilst it is rare for three galaxies to line up well enough for both of
the background sources to be strongly lensed \citep{GavazziEtal2008,CollettEtal2012a}, 
the large size of dark matter halos makes partial alignments -- such that they
act as perturbing weak lenses --  a near certainty. The large scale of the
perturbers means the external lensing perturbations can be approximated
by a quadrupole lens characterised only by external convergence and shear.
The external shear can potentially be recovered in the
mass-modelling of the lens, but the convergence is undetectable from the
image positions, shapes and relative fluxes; this is the famous {\emph{ 
mass sheet-degeneracy}} \citep[see e.g.][for details]{FalcoEtal1985}.

When inferring global quantities
(such as the cosmological parameters), combining the results from a
large number of independent objects will tend to reduce the uncertainty
due to external convergence \citep[\eg][]{Holz+Linder2005}, unless the lines of sight are not representative of the global population. If the line of sight selection function is not correctly modelled some residual systematic error will
remain even as the statistical uncertainty decreases. A selection function may be introduced either at the object
detection stage, or later on when making decisions about which objects
to study further. Large samples
of lenses are expected to be discovered in the next decade in
ground-based optical imaging surveys \citep{Oguri+Marshall2010}: both
the detectability of these lenses,  and the selection of sub-samples of
them for further observation, could be sensitive to the external
convergence.

Several attempts have been made to do better than simply constructing a
large sample of objects, and averaging over the resulting convergence
distribution. The weak lensing effect can be detected observationally 
by measuring the
small distortions it induces on the images of many background galaxies
in the field \citep[see e.g.][for a review]{Schneider2006}. 
\citet{NakajimaEtal2009} used deep HST imaging to infer $\kappax=0.17\pm0.06$ from weak lensing measurements; this measurement was used in the time delay distance
measurement of \citet{FadelyEtal2009}. A precise weak lensing estimate of $\kappax$ can only be made if the number density of of weakly lensed, measurable galaxies is high.

In order to make more precise estimates of the convergence at a given
sky position, we need more information.   There have been some attempts
to reconstruct the local and line of sight mass distribution in strong
lens fields in some detail, based on photometric and spectroscopic
surveys of the galaxies nearby. These efforts have focused on  detecting
galaxy groups, ascertaining the visible galaxies' membership of those
groups, and assigning mass to the group and galaxy halos in order to
predict the external shear apparently required by the strong lens
models. Surveys by
\citet{Fassnacht+Lubin2002,AugerEtal2007,WilliamsEtal2006,MomchevaEtal2006}
all found groups of galaxies both hosting, and lying  along the line of
sight to, known gravitational lenses. Initial estimates of the effect of
these structures suggested the importance of assigning mass to both
galaxy and group halos, and to allowing for significant uncertainty in
the centroids of group halos.

In order to propagate the effect of line-of-sight perturbations into 
uncertainties on time-delay distances we require the probability density function (PDF) for the external convergence
given our available data $\data$, $\Pr(\xkappa|\data)$
% PHIL's REVIEWS OF THE LITRATURE
\comments{\citet{GunnarssonEtal2006} investigated a related PDF,  for the
magnification, using a galaxy halo model. Applying empirical galaxy
scaling relations both to simulate mock catalogues and then to reconstruct
the line of sight mass distribution. While they did not explore highly
over-dense lines of sight, or the impact of groups and clusters, they
found that under their simplifying assumptions, the dispersion in
apparent source brightness due to lensing magnification could be reduced
by a factor of two.
%
\citet{WongEtal2011} estimated the PDF for the external shear,
$\Pr(\gammax|\data)$ from their survey data, running similar Monte Carlo
reconstructions of the mass in nine lens fields based on
\citeauthor{WilliamsEtal2006} and \citeauthor{MomchevaEtal2006}'s
photometric and spectroscopic measurements of galaxies close to the line
of sight. They then  compared the resulting predicted shear
distributions with the external shear demanded by the strong lens model.
Their reconstructions showed most of the shear to have been generated by
bright galaxies within 2~arcminutes of the lens, and that both line of
sight structures and the group of galaxies in each lens plane contribute
significant proportions of the shear. They found significant
discrepancies between the lens model shear and the Monte Carlo
predictions, but were unable to distinguish between the environment
modelling and the strong lens modelling as the cause of these
discrepancies.
%
\citet{KarpenkaEtal2012} used a halo model to predict deterministically 
the magnification due to
galaxies along the line of sight to type Ia supernovae, and inferred the
parameters of scaling relations between light and mass in these galaxies
from hundreds of supernova lightcone simultaneously. Following
\citet{JonssonEtal2010}, they subtract off the mean convergence to ensure that
the universal mass budget is balanced, and compare models where the halos are
all have either truncated singular isothermal profiles, or 
NFW profiles. Both groups were able to 
detect lensing effects in the supernova data this way.
}
% TOM's DISTILLING OF SAID REVIEWS.
Previous investigation of line-of-sight perturbations have focused on 
magnification and shear.\citet{GunnarssonEtal2006} used a galaxy halo model combined with simple scaling relations to reconstruct
the line of sight mass distribution and found that the dispersion in
apparent source brightness due to lensing magnification could be reduced
by a factor of two. Conversely \citet{KarpenkaEtal2012} used the brightness dispersion in type Ia supernovae to infer the parameters of both their galaxy halo model and mass-to-light scaling relation. Both \citeauthor{KarpenkaEtal2012} and \citet{JonssonEtal2010} were able to 
detect lensing effects in the supernova data this way.
%
\citet{WongEtal2011} estimated the PDF for the external shear,
$\Pr(\gammax|\data)$ from their survey data,
\citep{WilliamsEtal2006;MomchevaEtal2006} of galaxies close to the line
of sight. Comparing the predicted shear
distributions with the external shear demanded by the strong lens model 
they found significant
discrepancies, but were unable to distinguish between the environment
modelling and the strong lens modelling as the cause of these
discrepancies. \citeauthor{WongEtal2011} also found that both line of
sight structures and the group of galaxies in each lens plane contribute
significant proportions of the shear.

Large numerical simulations have been used to
predict, from first principles, the distribution of likely external
shears and convergences along various lines of sight, including those
containing strong lens systems.  \citet{Holder+Schechter2003} and
\citet{DalalEtal2005} carried out ray-tracing calculations in N-body
simulations to estimate the distribution of external shear values. 
\citet{HilbertEtal2009} performed similar ray-tracing experiments
through the Millennium Simulation \citep{SpringelEtal2005}, generating a
predicted $\kappax$ at every position in a simulated sky.
\citet{HilbertEtal2009} found that \MS lines of sight with strong lenses
were not biased towards high $\kappax$, although selection functions for
discovering or using particular samples of strong lenses were not taken
into account.

The \MS results have been used to analyse real observations as well. 
\citet{SuyuEtal2010} selected \MS lines of sight by their apparent
galaxy over-density, in a 45 arcsec radius aperture down to $i < 24.5$,
to match the observed over-density towards the time delay lens
B1608$+$656 \citep{FassnachtEtal2011}.  The resulting distribution of
$\kappax$ values from the ray-tracing was taken to be an estimate of 
$\Pr(\kappax)$, which was then marginalised over when inferring the time
delay distance in this system. A similar procedure was used in \citet{SuyuEtal2012b}.
% B1608$+$656 is indeed a high $\kappax$
% line of sight, but it does not reside in a massive cluster environment.
In a companion paper to this one, Greene et al. (submitted) investigate
improvements to this method by weighting the galaxy counts by galaxy
mass, and perpendicular distance to the line of sight, again using the
\MS mock catalogues and their associated $\kappax$ values to construct
$\Pr(\kappax|\data)$. For the most over-dense lines of sight the
Green et al. results  show a $\sim$20\% improvement over
number counts alone, but little improvement for less dense lines.

In this paper, we combine the halo model reconstruction approach of
\citeauthor{GunnarssonEtal2006}, \citeauthor{JonssonEtal2010}, 
\citeauthor{WongEtal2011} and others, with the
idea of calibrating to simulations from \citeauthor{SuyuEtal2010}.
Taking the $\kappax$ values from
\citeauthor{HilbertEtal2009}'s ray-tracing calculations as the ``truth''
that we have to recover, we use the \MS mock galaxy catalogues first to
{\it calibrate} the reconstruction, and account for unseen mass and
voids, and then to {\it test} the accuracy of the line of sight mass
reconstruction under this assumption. Probabilistically assigning mass
and redshift to every observed galaxy in a given observed field, we
generate Monte Carlo sample line of sight mass distributions, and so
construct the PDF $\Pr(\kappax|\data)$ for that field. We then emulate
the combination of many such PDFs to quantify the residual biases that
would be translated to the global (including cosmological) parameters.
In doing so, we aim to answer the following questions:

@@

@@

\citet{HilbertEtal2009} calculated the lensing convergence using the
``lightcone'' dark matter only 
output of the \MS, providing 1.7 degree square 
mock sky maps of this
quantity. In this process, the dark matter density from the simulation
was projected onto a set of lens planes at discrete redshifts,
adaptively gridded and smoothed, and then the second derivatives of the 
lensing potential required for the components of the magnification
matrix were computed using a multiple lens plane ray tracing algorithm.
The first order approximation to this algorithm \citep[equation 17
of][]{HilbertEtal2009} gives the total convergence at a given sky
position~$\bmath{\theta}$ as the simple summation of the surface density
in each lens plane, weighted by the inverse critical surface density
$\Sigma_{\rm cr}$ for that plane's redshift:
\be
\kappax(\bmath{\theta}) = \sum_i \frac{\Sigma_i(\bmath{\theta})}
                                     {\Sigma_{\rm cr}(z_i,\zs)}.
\ee
This approximation was found to be accurate to a few percent or better
in $\kappax$, even on 30 arcsecond scales.

@@

Cosmological dark matter simulations have shown that dark matter
halos are well-approximated by NFW profiles \citep{NFW}, with a density
profile given by
\be\label{eq:rhonfw}
\rho(r) = 
\frac{\rho_0}{\left(r/r_{s}\right)
\left(1+r/r_{s}\right)^{2}},
\ee
where $r_{s}$ is a characteristic radius of the cluster, representing where
the density slope transitions from $r^{-1}$ to $r^{-3}$. This radius is related
to the virial radius of the halo by $r_{s}~=~r_{200}/c$, where $r_{200}$ is the radius at 
which 
\be
\rho(r_{200}) = 200\rho_{{\mathrm{crit}}}(z), \text{ where } \rho_{\mathrm{crit}}(z) \equiv \frac{3H^{2}(z)}{8\pi G}
\ee
and $c$ is the concentration parameter, which can be estimated from the halo's mass,
using a mass--concentration relation. Typically more massive halos are less concentrated,
but there is some scatter; we use the relation of \citet{Neto2007} to estimate $c$
from the mass enclosed within $r_{200}$, which we denote as $M_{200}$. We find our results do not change if the \citet{MaccioEtal2008} Mass--concentration relation is used instead. The best fit relation of \citet{Neto2007} is given by
\be
c_{200} = 4.67 (M_{200}/10^{14} h^{-1} \Msun)^{-0.11}.
\ee

$\rho_0$ is calculated from the critical density and concentration parameter:
\be
\rho_0 = \rho_{\mathrm{crit}}(z)\frac{200}{3}\frac{c^{3}}{\ln(1+c)-c/(1+c)}.
\ee

If one integrates the density profile of an NFW profile out to infinite
radius, the total mass diverges. Similarly, if the universe is
homogeneously populated with NFW halos, the projected surface mass along
any line of sight will also be divergent.\footnote{At large radius,
$\Sigma_{\rm nfw} \propto R^{-2}$, but the differential number of halos
centred within an annulus of width $\dee R$ is given by $\dee N_{\rm
annulus} \propto R \dee R$, so  $\Sigma_{\rm total} \propto
\int_{0}^{\infty} R^{-1} \dee R$, which diverges logarithmically} Since
infinite mass is unphysical, the profile must be truncated at some
point. Several truncation profiles have been suggested
\citep[e.g][]{BMO}, but beyond several virial radii, the amount of
matter associated with a halo is likely to be low. In this work we
assume the truncated NFW profile
\be\label{eq:bmoprofile}
\rho(r) = 
\frac{\rho_{\rm NFW} (r)}{1+\left(r/r_{t}\right)^2},
\ee
which is the same as the NFW profile in the limit that the truncation
radius, $r_t$ goes to infinity. We use a truncation radius of five times
the virial radius, but our results are robust for any choice of $r_t>2
\times r_{200}$. The formula for the projected surface mass density 
$\Sigma(D_{od}\theta)$ for this
mass profile is given in \citet{BMO}.

Each halo in our catalogue then contributes to
the line of sight convergence by
\be
\label{eq:kappai}
\kappa_i =\Sigma_{i}/\Sigma_{\mathrm{cr}}(z_i,\zs),
\ee
where the critical surface density was defined in \Eref{eq:sigcrit}.
Following the first order approximation of \citet{HilbertEtal2009}
outlined in \Sref{sec:MS} above, we compute 
the total convergence from all the halos along the line
of sight using
\be 
\label{eq:kappasummu}
\kappah = \sum_{i} \kappa_i.
\ee
\notes{Using the Born approximation, we approximate light rays as straight.}

@@

While halos contribute a positive convergence to $\kappax$, voids contribute
a negative convergence. When a ray bundle passes through a  region of space
that is less dense than $\rho_{\rm mean}(z)$ the rays are de-focused\footnote{In
principle, the convergence caused by halos is also caused by the overdensity,
but we neglect this effect since halos are typically much more dense than the
intergalactic medium}. Whilst the  full ray-traced solution takes into account
the effect of voids, our reconstruction cannot -- the halo catalogue does not
include voids. \new{In principle the absence of galaxies implies something about
the presence a void, but we do not attempt to use this information}. Neglecting voids
would lead to a heavily biased estimate of $\kappax$. Moreover, the halos in 
the model only add mass to the background density, and that background density
already accounts for all the mass in the Universe! In order to make voids
negative relative to the mean density, we must subtract a significant smooth
mass component. The problem is that we do not have a good model for the
density structure of that smooth component.

The solution we present here is to calibrate the relationship between halo
convergence $\kappah$ and underlying convergence $\kappax$ (as calculated
during full ray tracing) using the simulations themselves.
The halo modelling \proceedure outlined in \Sref{sec:model:halos} allows 
us to estimate 
the convergence due to halos given observations of 
mass, redshift and position,
$\pr(\kappah|\mathcal{D})$, but what we are interested in is the total convergence
along the line of sight $\pr(\kappax|\mathcal{D})$. We can obtain this by
considering the expression:
\begin{equation}
\Pr(\kappax|\mathcal{D}) = \int \dee\kappah 
   \Pr(\kappax|\kappah) \Pr(\kappah|\mathcal{D})
\label{eq:kappaconv}   
\end{equation}
The first term in the integrand relates the convergence due to model halos,
$\kappah$, to the true convergence, $\kappax$. This conditional distribution
can be constructed from the Millennium Simulation catalogues, by computing
$\kappah$ for each selected line of sight using the true values of the halo
parameters,  and accumulating $\kappah$ and $\kappax$ along a large number of
such lines of
sight to  characterize $\Pr(\kappax|\kappah)$.
For any given line of sight we can then estimate
$\pr(\kappah|\data)$ as described above, and then
multiply it by $\Pr(\kappax|\kappah)$ and integrate out the intermediate
nuisance parameter $\kappah$. 
% We can do the marginalisation over $\kappah$
% by drawing samples from the two-dimensional grid and keeping only the
% $\kappa$ values, to form our final result, $\Pr(\kappa|\mathcal{D})$.

However, given an uncertain stellar mass to halo mass conversion, the PDF
$\pr(\kappah|\{\Mstar,z\})$ can become skewed relative to
$\pr(\kappah|\{\Mhalo,z\})$; this is because of the asymmetry in the
conversion from stellar mass through halo mass into $\kappah$. A long tail of
high halo masses gives rise to a tendency to overestimate $\kappah$.
If this shift is ignored, the resulting $\Pr(\kappax|\data)$ will be 
biased towards high values of $\kappax$.

We find that we can correct this bias to some extent by artificially 
suppressing the tails of $\pr(\kappah|\{\Mstar,z\})$. Let the median of
this PDF be denoted $\kappah^{\rm med}$. We again take a large number of
calibration sightlines from the \MS catalogues, but this time generate a mock
stellar mass catalogue for each one. We then reconstruct each lightcone in the
same way as we would an observed field,
estimate $\pr(\kappah|\{\Mstar,z\})$, and extract its median. Accumulating
these samples and their corresponding true values of $\kappax$, we can form the 
conditional distribution $\pr(\kappax|\kappah^{\rm med})$. 
Then, when inferring $\kappax$ from an observed photometric catalogue, we would
estimate $\pr(\kappah|\{\Mstar,z\}_{\rm obs})$ 
using the halo model, compute the median value 
$\kappa_{\rm h,obs}^{\rm med}$, and then use the following alternative
version of 
\Eref{eq:kappaconv} to infer $\kappax$: 
\bea
\Pr(\kappax|\mathcal{D}) &=& \int \dee\kappah^{\rm med} 
   \Pr(\kappax|\kappah^{\rm med}) \Pr(\kappah^{\rm med}|\mathcal{D}) \notag \\
                        &=& \Pr(\kappax|\kappa_{\rm h,obs}^{\rm med}).
\label{eq:calkappaconv}   
\eea
Here, we have replaced $\Pr(\kappah^{\rm med}|\mathcal{D})$ with a delta
function centred at the observed median value $\kappa_{\rm h,obs}$. 
The procedure of reducing the full PDF for $\kappah$ to just its median is
risky: the simulated lightcones that have median $\kappah$ values to match are
not guaranteed to provide an accurate estimate of $\kappax$ via this PDF
look-up. We proceed with caution, quantifying the biases and PDF widths in the
sections below.

>>>>

Along lines of sight containing strong gravitational lenses, the
perturbative effects of line of sight mass structure have been found to
be particularly important, with foreground and background structures
having a significant effect on the inferred lensing cross-section
\citep[\eg][]{WongEtal2012} and distance ratios
\citep[][]{DalalEtal2005}. Indeed, \citet{SuyuEtal2010} found that in
time delay lens cosmography the so-called ``external convergence''
$\kappax$ due to mass structures along the (unusually over-dense) line
of sight to the quadruply-imaged radio source B1608$+$656 was the
dominant source of uncertainty in their 5\% measurement of the time delay distance. 
Whilst it is rare for three galaxies to line up well enough for both of
the background sources to be strongly lensed \citep{GavazziEtal2008,CollettEtal2012a}, 
the large size of dark matter halos makes partial alignments -- such that they
act as perturbing weak lenses --  a near certainty. The large scale of the
perturbers means the external lensing perturbations can be approximated
by a quadrupole lens characterised only by external convergence and shear.
The external shear, $\gammax$ can potentially be recovered in the
mass-modelling of the lens, but the convergence is undetectable from the
image positions, shapes and relative fluxes; this is the famous {\emph{ 
mass-sheet degeneracy}} \citep[see e.g.][for details]{FalcoEtal1985}.

Large samples
of time-delay lenses are expected to be discovered in the next decade in
ground-based optical imaging surveys \citep{Oguri+Marshall2010}: both
the detectability of these lenses, and the selection of sub-samples of
them for further observation, could induce a selection function in external
convergence. External convergence must be included in analysis of time-delay
lenses else the inferred time-delay distance and cosmology be systematically biased.

Since the mass-sheet degeneracy prevents any estimate of $\kappax$ from the 
strong lens modelling additional information is required. Weak lensing measurements
can been used for rich lines of sight \citep{NakajimaEtal2009, FadelyEtal2009}, but 
for lines of sight with low galaxy density we must attempt to reconstruct the distribution
of mass along the line of sight. Attempts to reconstruct the mass distribution in strong
lens fields have focused on understanding the external shear which seems to be large in most strong lenses.
Surveys by \citet{Fassnacht+Lubin2002,AugerEtal2007,WilliamsEtal2006,MomchevaEtal2006}
all found groups of galaxies hosting or near to known gravitational lenses.
\citet{WongEtal2011} estimated $\Pr(\gammax)$ given the data of \citet{WilliamsEtal2006}
and \citet{MomchevaEtal2006} but found significant discrepancy between the predicted shear
distribution and the external shear demanded by the strong lens model. 
\citeauthor{WongEtal2011} also found that both line of
sight structures and the group of galaxies in each lens plane contribute
significant proportions of the shear.

For non-lens lines of sight magnification is a more useful quantity than convergence; \citet{GunnarssonEtal2006} used a galaxy halo model combined with simple scaling relations to reconstruct supernovae
lines of sight and found that the dispersion in
apparent source brightness due to lensing magnification could be reduced
by a factor of two. Conversely \citet{KarpenkaEtal2012} used the brightness dispersion in type Ia supernovae to infer the parameters of both their galaxy halo model and mass-to-light scaling relation. Both \citeauthor{KarpenkaEtal2012} and \citet{JonssonEtal2010} were able to 
detect lensing effects in the supernova data this way.

Large numerical simulations can also be used to estimate global convergence distributions.
 \citet{Holder+Schechter2003} and
\citet{DalalEtal2005} carried out ray-tracing calculations in N-body
simulations to estimate the distribution of external shear values. 
\citet{HilbertEtal2009} performed similar ray-tracing experiments
through the Millennium Simulation \citep{SpringelEtal2005}, generating a
predicted $\kappax$ at every position in a simulated sky.
\citet{HilbertEtal2009} found that after removing the matter on the strong lensing plane
\MS lines of sight with strong lenses
were not biased towards high $\kappax$, although selection functions for
discovering samples of strong lenses were not taken
into account.

The \MS results have been used to analyse real observations as well. 
\citet{SuyuEtal2010} selected \MS lines of sight by their apparent
galaxy over-density, in a 45 arcsec radius aperture down to $i < 24.5$,
to match the observed over-density towards the time delay lens
B1608$+$656 \citep{FassnachtEtal2011}.  The resulting distribution of
$\kappax$ values from the ray-tracing was taken to be an estimate of 
$\Pr(\kappax)$, which was then marginalised over when inferring the time
delay distance in this system. A similar procedure was used in \citet{SuyuEtal2012b}.
In a companion paper to this one, Greene et al. (submitted) investigate
improvements to this method by weighting the galaxy counts by galaxy
mass, and perpendicular distance to the line of sight, again using the
\MS mock catalogues and their associated $\kappax$ values to construct
$\Pr(\kappax|\data)$. For the most over-dense lines of sight the
Green et al. results  show a $\sim$20\% improvement over
number counts alone, but little improvement for less dense lines.

In this paper, we combine the halo model reconstruction approach of
\citeauthor{GunnarssonEtal2006}, \citeauthor{JonssonEtal2010}, 
\citeauthor{WongEtal2011} and others, with the
idea of calibrating to simulations from \citeauthor{SuyuEtal2010}
Taking the $\kappax$ values from
\citeauthor{HilbertEtal2009}'s ray-tracing calculations as the ``truth''
that we have to recover, we use the \MS mock galaxy catalogues first to
{\it calibrate} the reconstruction, and account for unseen mass and
voids, and then to {\it test} the accuracy of the line of sight mass
reconstruction under this assumption. Probabilistically assigning mass
and redshift to every observed galaxy in a given observed field, we
generate Monte Carlo sample line of sight mass distributions, and so
construct the PDF $\Pr(\kappax|\data)$ for that field. We then emulate
the combination of many such PDFs to quantify the residual biases that
would be translated to the global (including cosmological) parameters.
In doing so, we aim to answer the following questions:

@@

\comments{
When inferring global quantities
(such as the cosmological parameters), combining the results from a
large number of independent objects will tend to reduce the uncertainty
due to external convergence \citep[\eg][]{Holz+Linder2005}, unless the lines of sight are not representative of the global population. If the line of sight selection function is not correctly modelled some residual systematic error will
remain even as the statistical uncertainty decreases. A selection function may be introduced either at the object
detection stage, or later on when making decisions about which objects
to study further. Large samples
of lenses are expected to be discovered in the next decade in
ground-based optical imaging surveys \citep{Oguri+Marshall2010}: both
the detectability of these lenses,  and the selection of sub-samples of
them for further observation, could be sensitive to the external
convergence.

Several attempts have been made to do better than simply constructing a
large sample of objects, and averaging over the resulting convergence
distribution. The weak lensing effect can be detected observationally 
by measuring the
small distortions it induces on the images of many background galaxies
in the field \citep[see e.g.][for a review]{Schneider2006}. 
\citet{NakajimaEtal2009} used deep HST imaging to infer $\kappax=0.17\pm0.06$ from weak lensing measurements; this measurement was used in the time delay distance
measurement of \citet{FadelyEtal2009}. A precise weak lensing estimate of $\kappax$ can only be made if the number density of of weakly lensed, measurable galaxies is high.
}

\comments{In order to propagate the effect of line-of-sight perturbations into 
uncertainties on time-delay distances we require the probability density function (PDF) for the external convergence
given our available data $\data$, $\Pr(\xkappa|\data)$}
% PHIL's REVIEWS OF THE LITERATURE
\comments{\citet{GunnarssonEtal2006} investigated a related PDF,  for the
magnification, using a galaxy halo model. Applying empirical galaxy
scaling relations both to simulate mock catalogues and then to reconstruct
the line of sight mass distribution. While they did not explore highly
over-dense lines of sight, or the impact of groups and clusters, they
found that under their simplifying assumptions, the dispersion in
apparent source brightness due to lensing magnification could be reduced
by a factor of two.
%
\citet{WongEtal2011} estimated the PDF for the external shear,
$\Pr(\gammax|\data)$ from their survey data, running similar Monte Carlo
reconstructions of the mass in nine lens fields based on
\citeauthor{WilliamsEtal2006} and \citeauthor{MomchevaEtal2006}'s
photometric and spectroscopic measurements of galaxies close to the line
of sight. They then  compared the resulting predicted shear
distributions with the external shear demanded by the strong lens model.
Their reconstructions showed most of the shear to have been generated by
bright galaxies within 2~arcminutes of the lens, and that both line of
sight structures and the group of galaxies in each lens plane contribute
significant proportions of the shear. They found significant
discrepancies between the lens model shear and the Monte Carlo
predictions, but were unable to distinguish between the environment
modelling and the strong lens modelling as the cause of these
discrepancies.
%
\citet{KarpenkaEtal2012} used a halo model to predict deterministically 
the magnification due to
galaxies along the line of sight to type Ia supernovae, and inferred the
parameters of scaling relations between light and mass in these galaxies
from hundreds of supernova lightcone simultaneously. Following
\citet{JonssonEtal2010}, they subtract off the mean convergence to ensure that
the universal mass budget is balanced, and compare models where the halos are
all have either truncated singular isothermal profiles, or 
NFW profiles. Both groups were able to 
detect lensing effects in the supernova data this way.
}

@@

\citet{HilbertEtal2009} calculated the lensing convergence using the
``lightcone'' dark matter only 
output of the \MS, providing 1.7 degree square 
mock sky maps of this
quantity. In this process, the dark matter density from the simulation
was projected onto a set of lens planes at discrete redshifts,
adaptively gridded and smoothed, and then the second derivatives of the 
lensing potential required for the components of the magnification
matrix were computed using a multiple lens plane ray tracing algorithm.
The first order approximation to this algorithm \citep[equation 17
of][]{HilbertEtal2009} gives the total convergence at a given sky
position~$\bmath{\theta}$ as the simple summation of the surface density
in each lens plane, weighted by the inverse critical surface density
$\Sigma_{\rm cr}$ for that plane's redshift:
\be
\kappax(\bmath{\theta}) = \sum_i \frac{\Sigma_i(\bmath{\theta})}
                                     {\Sigma_{\rm cr}(z_i,\zs)}.
\ee
\citet{HilbertEtal2009} found this approximation was found to be accurate to a few percent or better
in $\kappax$, even on 30 arcsecond scales.

@@

Cosmological dark matter simulations have shown that dark matter
halos are well-approximated by NFW profiles

\citep{NFW}, with a density
profile given by
\be\label{eq:rhonfw}
\rho(r) = 
\frac{\rho_0}{\left(r/r_{s}\right)
\left(1+r/r_{s}\right)^{2}},
\ee
where $r_{s}$ is a characteristic radius of the cluster, representing where
the density slope transitions from $r^{-1}$ to $r^{-3}$. This radius is related
to the virial radius of the halo by $r_{s}~=~r_{200}/c$, where $c$ is the concentration parameter, which can be estimated from the halo's mass,
using a mass--concentration relation. Typically more massive halos are less concentrated,
but there is some scatter; we use the relation of \citet{Neto2007} to estimate $c$
from the mass enclosed within $r_{200}$, which we denote as $M_{200}$. We find our results do not change if the \citet{MaccioEtal2008} Mass--concentration relation is used instead.

If one integrates the density profile of an NFW profile out to infinite
radius, the total mass diverges. Similarly, if the universe is
homogeneously populated with NFW halos, the projected surface mass along
any line of sight will also be divergent.\footnote{At large radius,
$\Sigma_{\rm nfw} \propto R^{-2}$, but the differential number of halos
centred within an annulus of width $\dee R$ is given by $\dee N_{\rm
annulus} \propto R \dee R$, so  $\Sigma_{\rm total} \propto
\int_{0}^{\infty} R^{-1} \dee R$, which diverges logarithmically} Since
infinite mass is unphysical, the profile must be truncated at some
point. Several truncation profiles have been suggested
\citep[e.g][]{BMO}, but beyond several virial radii, the amount of
matter associated with a halo is likely to be low. In this work we
assume the truncated NFW profile
\be\label{eq:bmoprofile}
\rho(r) = 
\frac{\rho_{\rm NFW} (r)}{1+\left(r/r_{t}\right)^2},
\ee
which is the same as the NFW profile in the limit that the truncation
radius, $r_t$ goes to infinity; the shear and convergence from such a
profile are derived in \citet{BMO}.
 We use a truncation radius of five times
the virial radius, but our results are robust for any choice of $r_t>2
\times r_{200}$.

Each halo in our catalogue then contributes to
the line of sight convergence by
\be
\label{eq:kappai}
\kappa_i =\Sigma_{i}/\Sigma_{\mathrm{cr}}(z_i,\zs),
\ee
where the critical surface density was defined in \Eref{eq:sigcrit}.
Following the first order approximation of \citet{HilbertEtal2009}
outlined in \Sref{sec:MS} above, we compute 
the total convergence from all the halos along the line
of sight using
\be 
\label{eq:kappasummu}
\kappah = \sum_{i} \kappa_i.
\ee

@@

Our halo model accounts for the convergence contribution of density perturbations that can
be associated with light. Filaments and dark substructures contribute additional convergence
whilst voids contribute a divergence; these structures must be included to produce an unbiased estimate of $\kappax$
in principle the absence of galaxies implies something about
the presence a void, but we do not currently attempt to use this information. Neglecting voids
would lead to a heavily biased estimate of $\kappax$, since the halo model's $\kappah$
can only be positive. It is difficult to correct for unseen structure problem
since we do not have a good model for its density structure.

The solution we present here is to calibrate the relationship between halo
convergence $\kappah$ and underlying convergence $\kappax$ (as calculated
during full ray tracing) using the simulations themselves.
The halo modelling procedure outlined in \Sref{sec:model:halos} allows 
us to estimate 
the convergence due to halos given observations of 
mass, redshift and position,
$\pr(\kappah|\mathcal{D})$, but what we are interested in is the total convergence
along the line of sight $\pr(\kappax|\mathcal{D})$. We can obtain this by
considering the expression:
\begin{equation}
\Pr(\kappax|\mathcal{D}) = \int \dee\kappah 
   \Pr(\kappax|\kappah,\mathcal{D}) \Pr(\kappah|\mathcal{D})
\label{eq:kappaconv}   
\end{equation}
The first term in the integrand relates the convergence due to model halos,
$\kappah$, to the true convergence, $\kappax$. This conditional distribution
can be constructed from the Millennium Simulation catalogues, by computing
$\kappah$ for each selected line and accumulating $\kappah$ and $\kappax$ along a large number of
such lines of
sight to  characterize $\Pr(\kappax|\kappah)$.
For any given line of sight we can then estimate
$\pr(\kappah|\data)$ as described above, and then
multiply it by $\Pr(\kappax|\kappah)$ and integrate out the intermediate
parameter $\kappah$.

However, given an uncertain stellar mass to halo mass conversion,
$\pr(\kappah|\mathcal{D}_1)$ can be shifted relative to
$\pr(\kappah|\mathcal{D}_2)$ given perfect knowledge of the halo mass; this is because of the asymmetry in the
conversion from stellar mass through halo mass into $\kappah$. A long tail of
high halo masses gives rise to a tendency to overestimate $\kappah$.
If this shift is ignored, the resulting $\Pr(\kappax|\data)$ will be 
biased towards high values of $\kappax$ if $\Pr(\kappax|\kappah)$ is constructed from lines
of sight with perfect knowledge of the halo mass. Ideally one wants to find the $\kappax$ distribution
of all calibration lines of sight with a $\pr(\kappah|\data)$ that is identical to $\pr(\kappah|\data)$ for the real line of sight
{\it given the same data quality}; this is impossible given a finite number of calibration lines of sight.
We compromise on the similarity of our real line of sight to the calibration lines of sight by 
selecting considering only the median of the $\pr(\kappah|\data)$ distributions.

We  take a large number of
calibration sightlines from the \MS catalogues and generate a mock
stellar mass catalogue for each one. We then reconstruct each lightcone in the
same way as we would an observed field,
estimate $\pr(\kappah|\data)$, and extract its median. Accumulating
these samples and their corresponding true values of $\kappax$, we can form the 
conditional distribution $\pr(\kappax|\kappah^{\rm med})$. 
Then, when inferring $\kappax$ from an observed photometric catalogue, we 
estimate $\pr(\kappah|\data)$ 
using the halo model, compute the median value 
$\kappa_{\rm h,obs}^{\rm med}$, and then use an alternative
version of 
\Eref{eq:kappaconv} to infer $\kappax$: 
\bea
\Pr(\kappax|\kappah^{\rm med},\mathcal{D}) &=& \int \dee\kappah^{\rm med} 
   \Pr(\kappax|\kappah^{\rm med},\mathcal{D}) \Pr(\kappah^{\rm med}|\mathcal{D}) \notag \\
\label{eq:calkappaconv}   
\eea
Because we only have a finite number of calibration lines of sight, we use all calibration
lines of sight with $kappah_{\rm calibration}^{\rm med} = kappah^{\rm med}\pm0.003$ to form
$\Pr(\kappax|\kappah^{\rm med},\mathcal{D})$, with each calibration line weighted
equally.

The procedure of reducing the full PDF for $\kappah$ to its median increases the 
importance of the simulation choice: the simulation needs to be a similar as possible
to the real universe or there is a possibility of systematic errors. The reader
should keep in mind that we have only performed this analysis using the \MS.
